---
title: "MOPA Dominance Rank"
author: 
date: 
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Load packages
```{r, include=FALSE, echo=FALSE}
library(dplyr)
library(tidyr)
library(reshape2)
library(domstruc) #install_github("danm0nster/domstruc")
library(igraph)
library(tidyverse)
library(here)


```

Rank detection code adapted from Dr.Annemarie van der Marel


#2021
## import data
```{r data import, echo=FALSE}
behave <- read_csv("2021_interactions_raw.csv") 
unique(behave$date)


# period file
obs.day.key <-  read_csv("2021_timeline_fieldwork.csv")
bin.dates <-  read_csv("2021_3daybins.csv") 
# capture file to obtain bird list
birdIDs<- read_csv("2021_birdIDs.csv") %>%
  dplyr::select(mark_id3) %>%
  filter(mark_id3!="") # all 20 birds
birdIDs_removal1 <- birdIDs %>% filter(mark_id3!="BBB")
birdIDs_removal2 <- birdIDs %>% filter(mark_id3!="GPG")
birdIDs_removal3 <- birdIDs %>% filter(mark_id3!="OPP")

```


```{r}


### change bird names for actor and subject, if using all data
 #(mark_id1 = mark_id2)
 #we then have to allocate a line for each of these birds with 0's to make the matrix complete

# GBB = NPB (mark_id1 = mark_id2)
behave$actor[behave$actor=="GBB"] <- "NPB"
behave$subject[behave$subject=="GBB"] <- "NPB"

# BOO = POO (mark_id1 = mark_id2)
behave$actor[behave$actor=="BOO"] <- "POO"
behave$subject[behave$subject=="BOO"] <- "POO"

# NPB = PPB (mark_id2 = mark_id3)
behave$actor[behave$actor=="NPB"] <- "PPB"
behave$subject[behave$subject=="NPB"] <- "PPB"

# check
unique(behave$actor)
unique(behave$subject)
```

## get data ready 

to compare hierarchy over time include aggression at feeder

From 2021-04-24, aggression at feeders separate but to compare with beginning add aggression at feeders with the rest
```{r}
# for feeder 
feed_agg <- c("displace_feed", "crowd_feed", "spar_feed")

# non-feeder aggression
nonfeed_agg <- c("displace", "crowd", "spar")

# all aggression (both feeders and nonfeeders)
behave_agg <- behave %>%
  separate(behavior, c("behavior", "type"), sep="_") # all aggression at both feeders and away from feeders

unique(behave_agg$behavior)
unique(behave_agg$date)

```

##check per perturbation period
```{r}
# per 3 day chunks 
glimpse(obs.day.key)

bin<- bin.dates %>%
  dplyr::select(date, bin) 


bin$bin <- as.numeric(bin$bin)

# all aggression -> both at feeder and nonfeeder
behave_totalagg <- left_join(behave, bin, by="date") 


unique(behave_totalagg$bin)
unique(behave_totalagg$date)

#excluding aggression at feeder
behave_nonfeed <- left_join(behave, bin, by="date")  
unique(behave_nonfeed$bin)
unique(behave_nonfeed$date)


# 
# behave_nonfeed <- left_join(behave, bin, by="date") %>%
#   dplyr::filter(bin==23 | bin==25) # excluding aggression at feeder
# unique(behave_nonfeed$bin)
# unique(behave_nonfeed$date)

```

## Helper functions & data
```{r echo=FALSE}

#make matrix function (from dataframe, 1st col with row names)
matrix.please<-function(x) {
  m<-as.matrix(x[,-1])
  rownames(m)<-x[,1]
  m
}

##list of all valid color combinations
 
bird.list <- sort(unique(birdIDs$mark_id3)) 
length(bird.list)


# list of all possible dyads
dyad.list <- expand.grid(bird.list, bird.list) #head(dyad.list)
names(dyad.list) <- c("actor", "subject")
dyad.list <- subset(dyad.list, actor!=subject)
dyad.list$dyadID <- paste(dyad.list$actor, dyad.list$subject, sep="-")

# list of days with observations
obsday.list <- as.character(unique(behave$date)) #only days with observations
obsday.list.wday <- as.data.frame(obsday.list)
colnames(obsday.list.wday) <- "date"
obsday.list.wday$nday <- seq(1:length(obsday.list.wday$date))


```


## bin level more conservative: removing duplicates by crowds and displacements combined
```{r Dominance using crowds and displacements more conservative, echo=FALSE}
#### Remove duplicates for displacements and crowds ####


# filter displacement & crowd data from all data
aggDC <- behave_totalagg %>%
  subset(behavior=="displace" | behavior=="crowd" | behavior=="displace_feed" | behavior=="crowd_feed") %>% 
  dplyr::select(sessionKEY, session_start_timeStamp, bin, date, time, actor, subject, behavior)

aggDC$date <- as.Date(aggDC$date)

#add  May 15 into bin 2
aggDC$bin[aggDC$date == as.Date("2021-05-15")] <- 2

#PPB doesnt show up in the study period (5-15 to 5-18 so remove from birdlist)
bird.list_ppb_removed <-  bird.list[bird.list != "PPB"]

str(aggDC)
length(aggDC$sessionKEY)

# keep only observations where the actor is in the list of bird IDs
aggDC.goodID <- subset(aggDC, actor %in% bird.list_ppb_removed)

# keep only observations where the actor (found above) and the subject are both in the list of bird IDs
aggDC.goodID <- subset(aggDC.goodID, subject %in% bird.list_ppb_removed)


## for total agonistic events (crowds and displacements combined)
#finds n CD and crowds by sessionKEY, date, time, actor, subject, behavior
dyad.totalagg.key <- aggDC.goodID %>% 
  dplyr::group_by(sessionKEY, bin, date, time,actor, subject) %>%  # removed , behavior
  tally() # counts behaviors that were observed within the same minute

#finds max agg per actor by date, time, and behavior type. Return just that summary
dyad.totalagg.maxkey <- dyad.totalagg.key %>% 
  group_by(actor, subject,bin,  date, time) %>% # removed behavior, behavior
  slice(which.max(n)) %>% # n.aggXaggXkey if summarise, n if tally
  ungroup()
head(dyad.totalagg.maxkey)

#finds n disps per actor by date for summarized/trimmed data
dyad.totalaggXbin <- dyad.totalagg.maxkey %>% 
  group_by(actor, subject, bin) %>% # removed behavior to get total aggression
  summarise(n.aggXbin=sum(n)) %>% 
  ungroup()
head(dyad.totalaggXbin)
sum(dyad.totalaggXbin$n.aggXbin)

#Here we are interested in total number of interactions instead of chronological #order of interactions. If you require order of interactions, then next steps will #be different from the following setup as you will have to add the behaviors with #n>1 as separate rows. 

## Summarize displacements & crowds combined --> total aggression

head(dyad.totalaggXbin) # crowds and displacement combined from the start, which is more conservative
str(dyad.totalaggXbin)

# check 
unique(dyad.totalaggXbin$actor)
unique(dyad.totalaggXbin$subject)

sum(dyad.totalaggXbin$n.aggXbin)

# summarize total aggression events per day
sum.aggDCXbin_conservative <-dyad.totalaggXbin %>%
  group_by(bin) %>%
  summarize(n.agg = sum(n.aggXbin), n.dyads=length(actor))
head(sum.aggDCXbin_conservative)

n.aggDC.conservative <- sum(sum.aggDCXbin_conservative$n.agg)


#### DOMINANCE RANK BY bin####

head(dyad.totalaggXbin)

# make empty data frame to put data in
ec_rank <- data.frame(run.code=character(),
                      bird.list_ppb_removed=character(), 
                      dom_ec=numeric(),
                      between=numeric())

# run = 1

test <- subset(dyad.totalaggXbin, bin == 2)


start.time <- Sys.time()

dyad.totalaggXbin <- dyad.totalaggXbin %>% filter(!is.na(bin))


loopby<-unique(dyad.totalaggXbin$bin)

#i=1

for(i in seq_along(unique(dyad.totalaggXbin$bin))){
  run.code <- loopby[i]
  print(run.code)
  run.data <- subset(dyad.totalaggXbin, bin==run.code)
  
# number of dyads correct?
  #length(run.data$actor)
  #check <- sum.aggXday %>% filter(date==run.code)
  #head(check)
  
allowed <- union(unique(run.data$actor), unique(run.data$subject))

# filter df1 to rows whose actor AND subject are in df2 somewhere
ddyad.list_filtered <- dyad.list %>%
  filter(actor %in% allowed & subject %in% allowed)
 

bird.list_ppb_removed_filtered <- bird.list_ppb_removed[bird.list_ppb_removed %in% allowed & !is.na(bird.list_ppb_removed)]


  data_alldyads <- merge(ddyad.list_filtered, run.data, all.x=TRUE,
                       by=c("actor", "subject")) 
  
  #head(data_alldyads)
  #length(unique(data_alldyads$dyadID)) # with 21 birds should be 420 and with 22 462 dyads
  
  data_alldyads$date<-run.code
  
 #fill newly-merged data with 0's where no interactions
 data_alldyads[is.na(data_alldyads)] <- 0
  
   #print a check
  check <- length(data_alldyads$actor)
  print(check)

# matrix  
  mx <- reshape2::dcast(data_alldyads, actor~subject, value.var="n.aggXbin") #head(ref.behavior1.mx)
  mx[is.na(mx)] <- 0 #for linearity measure, matrix needs to be fully filled, no NAs
  mx <- matrix.please(mx)
  
  ## dominance
dom_ec <- dom_ec(mx)

## Betweenness centrality
# network
  graph <- graph_from_adjacency_matrix(mx, "directed", weighted=T, diag=F)

# betweenness centrality
between <- betweenness(graph, 
                          directed = TRUE, 
                          weights = E(graph)$weight)


ids <- rownames(mx)
# combine in dataframe
pool <- cbind.data.frame(run.code, 
                         bird.list_ppb_removed_filtered,
                         dom_ec,
                         between)

  ec_rank<-rbind(ec_rank, pool )

}

end.time <- Sys.time()

#Time to run:
end.time - start.time

# check 
head(ec_rank)
length(unique(ec_rank$run.code))
check<-ec_rank %>% group_by(run.code) %>% summarize(length(bird.list_ppb_removed_filtered))

ec_rankXbin<-ec_rank %>%
  rename(bin=run.code, id=bird.list_ppb_removed_filtered)

# higher rank is higher position in the hierarchy?
head(ec_rankXbin)


ec_rankXbin <- arrange(ec_rankXbin, bin, dom_ec)

MOPA_bin_rank_2021_test <- ec_rankXbin %>%
  group_by(bin) %>%
  mutate(rank = dense_rank(dom_ec)) %>%
  ungroup()


MOPA_bin_rank_2021_test <- mutate(MOPA_bin_rank_2021_test, power = (1- dom_ec))

#write.csv(MOPA_bin_rank_2021_test,  "MOPA_bin_rank_2021_test.csv", row.names = FALSE)
```


#2022G1
## import data
```{r data import, echo=FALSE}
#import data
behave <- read_csv("2022_interactions_group1.csv") 

unique(behave$date)


# period file
bin.dates <-  read_csv("2022_3daybins.csv") #%>%
  

# capture file to obtain bird list
birdIDs<- read_csv("2022_birdIDs.csv") %>%
  dplyr::select(mark_id22) %>%
  filter(mark_id22!="") 


```

## get data ready 

to compare hierarchy over time include aggression at feeder

```{r}
# for feeder 
feed_agg <- c("displace_feed", "crowd_feed", "spar_feed")

# non-feeder aggression
nonfeed_agg <- c("displace", "crowd", "spar")

# all aggression (both feeders and nonfeeders)
behave_agg <- behave %>%
  separate(behavior, c("behavior", "type"), sep="_") # all aggression at both feeders and away from feeders

unique(behave_agg$behavior)
unique(behave_agg$date)

```


##check per perturbation period
```{r}
# per 3 day chunks 

bin<- bin.dates %>%
  dplyr::select(date, bin) 


bin$bin <- as.numeric(bin$bin)
bin$date <- as.Date(bin$date, format = "%m/%d/%Y")
#use to check all aggression
behave_totalagg <- left_join(behave, bin, by="date") 

unique(behave_totalagg$bin)
unique(behave_totalagg$date)

#use if want to check only nonfeeder aggression
behave_nonfeed <- left_join(behave, bin, by="date") 
unique(behave_nonfeed$bin)
unique(behave_nonfeed$date)


# 
# behave_nonfeed <- left_join(behave, bin, by="date") %>%
#   dplyr::filter(bin==23 | bin==25) # excluding aggression at feeder
# unique(behave_nonfeed$bin)
# unique(behave_nonfeed$date)

```


## Helper functions & data
```{r echo=FALSE}

#make matrix function (from dataframe, 1st col with row names)
matrix.please<-function(x) {
  m<-as.matrix(x[,-1])
  rownames(m)<-x[,1]
  m
}

##list of all valid color combinations
bird.list <- sort(unique(birdIDs$mark_id22)) 
length(bird.list)

# ID's of removed birds removal


# list of all possible dyads
dyad.list <- expand.grid(bird.list, bird.list) #head(dyad.list)
names(dyad.list) <- c("actor", "subject")
dyad.list <- subset(dyad.list, actor!=subject)
dyad.list$dyadID <- paste(dyad.list$actor, dyad.list$subject, sep="-")

# list of days with observations
obsday.list <- as.character(unique(behave$date)) #only days with observations
obsday.list.wday <- as.data.frame(obsday.list)
colnames(obsday.list.wday) <- "date"
obsday.list.wday$nday <- seq(1:length(obsday.list.wday$date))


```


## bin level more conservative: removing duplicates by crowds and displacements combined
```{r Dominance using crowds and displacements more conservative, echo=FALSE}
#### Remove duplicates for displacements and crowds ####

# filter displacement & crowd data from all data
aggDC <- behave_totalagg %>%
  subset(behavior=="displace" | behavior=="crowd" | behavior=="displace_feed" | behavior=="crowd_feed") %>% 
  dplyr::select(sessionKEY, session_start_timeStamp, bin, date, time, actor, subject, behavior)

str(aggDC)
length(aggDC$sessionKEY)

# keep only observations where the actor is in the list of bird IDs
aggDC.goodID <- subset(aggDC, actor %in% bird.list)

# keep only observations where the actor (found above) and the subject are both in the list of bird IDs
aggDC.goodID <- subset(aggDC.goodID, subject %in% bird.list)


## for total agonistic events (crowds and displacements combined)
#finds n CD and crowds by sessionKEY, date, time, actor, subject, behavior
dyad.totalagg.key <- aggDC.goodID %>% 
  dplyr::group_by(sessionKEY, bin, date, time,actor, subject) %>%  # removed , behavior
  tally() # counts behaviors that were observed within the same minute

#finds max agg per actor by date, time, and behavior type. Return just that summary
dyad.totalagg.maxkey <- dyad.totalagg.key %>% 
  group_by(actor, subject,bin,  date, time) %>% # removed behavior, behavior
  slice(which.max(n)) %>% # n.aggXaggXkey if summarise, n if tally
  ungroup()
head(dyad.totalagg.maxkey)

#finds n disps per actor by date for summarized/trimmed data
dyad.totalaggXbin <- dyad.totalagg.maxkey %>% 
  group_by(actor, subject, bin) %>% # removed behavior to get total aggression
  summarise(n.aggXbin=sum(n)) %>% 
  ungroup()
head(dyad.totalaggXbin)
sum(dyad.totalaggXbin$n.aggXbin)

#Here we are interested in total number of interactions instead of chronological #order of interactions. If you require order of interactions, then next steps will #be different from the following setup as you will have to add the behaviors with #n>1 as separate rows. 

## Summarize displacements & crowds combined --> total aggression

head(dyad.totalaggXbin) # crowds and displacement combined from the start, which is more conservative
str(dyad.totalaggXbin)

# check 
unique(dyad.totalaggXbin$actor)
unique(dyad.totalaggXbin$subject)

sum(dyad.totalaggXbin$n.aggXbin)

# summarize total aggression events per day
sum.aggDCXbin_conservative <-dyad.totalaggXbin %>%
  group_by(bin) %>%
  summarize(n.agg = sum(n.aggXbin), n.dyads=length(actor))
head(sum.aggDCXbin_conservative)

n.aggDC.conservative <- sum(sum.aggDCXbin_conservative$n.agg)


#### DOMINANCE RANK BY bin####

head(dyad.totalaggXbin)

# make empty data frame to put data in
ec_rank <- data.frame(run.code=character(),
                      bird.list=character(), 
                      dom_ec=numeric(),
                      between=numeric())

# run = 1

start.time <- Sys.time()

dyad.totalaggXbin <- dyad.totalaggXbin %>% filter(!is.na(bin))


loopby<-unique(dyad.totalaggXbin$bin)

#i=1

for(i in seq_along(unique(dyad.totalaggXbin$bin))){
  run.code <- loopby[i]
  print(run.code)
  run.data <- subset(dyad.totalaggXbin, bin==run.code)
  
# number of dyads correct?
  #length(run.data$actor)
  #check <- sum.aggXday %>% filter(date==run.code)
  #head(check)
  
allowed <- union(unique(run.data$actor), unique(run.data$subject))

# filter df1 to rows whose actor AND subject are in df2 somewhere
ddyad.list_filtered <- dyad.list %>%
  filter(actor %in% allowed & subject %in% allowed)
 

bird.list_filtered <- bird.list[bird.list %in% allowed & !is.na(bird.list)]


  data_alldyads <- merge(ddyad.list_filtered, run.data, all.x=TRUE,
                       by=c("actor", "subject")) 
  
  #head(data_alldyads)
  #length(unique(data_alldyads$dyadID)) # with 21 birds should be 420 and with 22 462 dyads
  
  data_alldyads$date<-run.code
  
 #fill newly-merged data with 0's where no interactions
 data_alldyads[is.na(data_alldyads)] <- 0
  
   #print a check
  check <- length(data_alldyads$actor)
  print(check)

# matrix  
  mx <- reshape2::dcast(data_alldyads, actor~subject, value.var="n.aggXbin") #head(ref.behavior1.mx)
  mx[is.na(mx)] <- 0 #for linearity measure, matrix needs to be fully filled, no NAs
  mx <- matrix.please(mx)
  
  ## dominance
dom_ec <- dom_ec(mx)

## Betweenness centrality
# network
  graph <- graph_from_adjacency_matrix(mx, "directed", weighted=T, diag=F)

# betweenness centrality
between <- betweenness(graph, 
                          directed = TRUE, 
                          weights = E(graph)$weight)



# combine in dataframe
pool <- cbind.data.frame(run.code, 
                         bird.list_filtered,
                         dom_ec,
                         between)

  ec_rank<-rbind(ec_rank, pool )

}

end.time <- Sys.time()

#Time to run:
end.time - start.time

# check 
head(ec_rank)
length(unique(ec_rank$run.code))
check<-ec_rank %>% group_by(run.code) %>% summarize(length(bird.list_filtered))

ec_rankXbin<-ec_rank %>%
  rename(bin=run.code, id=bird.list_filtered)

# higher rank is higher position in the hierarchy
head(ec_rankXbin)


ec_rankXbin <- arrange(ec_rankXbin, bin, dom_ec)

MOPA_bin_rank_2022G1 <- ec_rankXbin %>%
  group_by(bin) %>%
  mutate(rank = dense_rank(dom_ec)) %>%
  ungroup()


#write.csv(MOPA_bin_rank_2022G1,"MOPA_bin_rank_2022G1.csv", row.names = FALSE)


```

#2022G2
## import data
```{r data import, echo=FALSE}
#import data
behave <- read_csv("2022_interactions_group2.csv") 

unique(behave$date)


# period file
bin.dates <-  read_csv("2022_3daybins.csv") #%>%
  

# capture file to obtain bird list
birdIDs<- read_csv("2022_birdIDs.csv") %>%
  dplyr::select(mark_id22) %>%
  filter(mark_id22!="") 


```

## get data ready 

to compare hierarchy over time include aggression at feeder

```{r}
# for feeder 
feed_agg <- c("displace_feed", "crowd_feed", "spar_feed")

# non-feeder aggression
nonfeed_agg <- c("displace", "crowd", "spar")

# all aggression (both feeders and nonfeeders)
behave_agg <- behave %>%
  separate(behavior, c("behavior", "type"), sep="_") # all aggression at both feeders and away from feeders

unique(behave_agg$behavior)
unique(behave_agg$date)

```


##check per perturbation period
```{r}
# per 3 day chunks 

bin<- bin.dates %>%
  dplyr::select(date, bin) 


bin$bin <- as.numeric(bin$bin)
bin$date <- as.Date(bin$date, format = "%m/%d/%Y")
#use to check all aggression
behave_totalagg <- left_join(behave, bin, by="date") 

unique(behave_totalagg$bin)
unique(behave_totalagg$date)

#use if want to check only nonfeeder aggression
behave_nonfeed <- left_join(behave, bin, by="date") 
unique(behave_nonfeed$bin)
unique(behave_nonfeed$date)


# 
# behave_nonfeed <- left_join(behave, bin, by="date") %>%
#   dplyr::filter(bin==23 | bin==25) # excluding aggression at feeder
# unique(behave_nonfeed$bin)
# unique(behave_nonfeed$date)

```


## Helper functions & data
```{r echo=FALSE}

#make matrix function (from dataframe, 1st col with row names)
matrix.please<-function(x) {
  m<-as.matrix(x[,-1])
  rownames(m)<-x[,1]
  m
}

##list of all valid color combinations
bird.list <- sort(unique(birdIDs$mark_id22)) 
length(bird.list)

# ID's of removed birds removal


# list of all possible dyads
dyad.list <- expand.grid(bird.list, bird.list) #head(dyad.list)
names(dyad.list) <- c("actor", "subject")
dyad.list <- subset(dyad.list, actor!=subject)
dyad.list$dyadID <- paste(dyad.list$actor, dyad.list$subject, sep="-")

# list of days with observations
obsday.list <- as.character(unique(behave$date)) #only days with observations
obsday.list.wday <- as.data.frame(obsday.list)
colnames(obsday.list.wday) <- "date"
obsday.list.wday$nday <- seq(1:length(obsday.list.wday$date))


```


## bin level more conservative: removing duplicates by crowds and displacements combined
```{r Dominance using crowds and displacements more conservative, echo=FALSE}
#### Remove duplicates for displacements and crowds ####

# filter displacement & crowd data from all data
aggDC <- behave_totalagg %>%
  subset(behavior=="displace" | behavior=="crowd" | behavior=="displace_feed" | behavior=="crowd_feed") %>% 
  dplyr::select(sessionKEY, session_start_timeStamp, bin, date, time, actor, subject, behavior)

str(aggDC)
length(aggDC$sessionKEY)

# keep only observations where the actor is in the list of bird IDs
aggDC.goodID <- subset(aggDC, actor %in% bird.list)

# keep only observations where the actor (found above) and the subject are both in the list of bird IDs
aggDC.goodID <- subset(aggDC.goodID, subject %in% bird.list)


## for total agonistic events (crowds and displacements combined)
#finds n CD and crowds by sessionKEY, date, time, actor, subject, behavior
dyad.totalagg.key <- aggDC.goodID %>% 
  dplyr::group_by(sessionKEY, bin, date, time,actor, subject) %>%  # removed , behavior
  tally() # counts behaviors that were observed within the same minute

#finds max agg per actor by date, time, and behavior type. Return just that summary
dyad.totalagg.maxkey <- dyad.totalagg.key %>% 
  group_by(actor, subject,bin,  date, time) %>% # removed behavior, behavior
  slice(which.max(n)) %>% # n.aggXaggXkey if summarise, n if tally
  ungroup()
head(dyad.totalagg.maxkey)

#finds n disps per actor by date for summarized/trimmed data
dyad.totalaggXbin <- dyad.totalagg.maxkey %>% 
  group_by(actor, subject, bin) %>% # removed behavior to get total aggression
  summarise(n.aggXbin=sum(n)) %>% 
  ungroup()
head(dyad.totalaggXbin)
sum(dyad.totalaggXbin$n.aggXbin)

#Here we are interested in total number of interactions instead of chronological #order of interactions. If you require order of interactions, then next steps will #be different from the following setup as you will have to add the behaviors with #n>1 as separate rows. 

## Summarize displacements & crowds combined --> total aggression

head(dyad.totalaggXbin) # crowds and displacement combined from the start, which is more conservative
str(dyad.totalaggXbin)

# check 
unique(dyad.totalaggXbin$actor)
unique(dyad.totalaggXbin$subject)

sum(dyad.totalaggXbin$n.aggXbin)

# summarize total aggression events per day
sum.aggDCXbin_conservative <-dyad.totalaggXbin %>%
  group_by(bin) %>%
  summarize(n.agg = sum(n.aggXbin), n.dyads=length(actor))
head(sum.aggDCXbin_conservative)

n.aggDC.conservative <- sum(sum.aggDCXbin_conservative$n.agg)


#### DOMINANCE RANK BY bin####

head(dyad.totalaggXbin)

# make empty data frame to put data in
ec_rank <- data.frame(run.code=character(),
                      bird.list=character(), 
                      dom_ec=numeric(),
                      between=numeric())

# run = 1

start.time <- Sys.time()

dyad.totalaggXbin <- dyad.totalaggXbin %>% filter(!is.na(bin))


loopby<-unique(dyad.totalaggXbin$bin)

#i=1

for(i in seq_along(unique(dyad.totalaggXbin$bin))){
  run.code <- loopby[i]
  print(run.code)
  run.data <- subset(dyad.totalaggXbin, bin==run.code)
  
# number of dyads correct?
  #length(run.data$actor)
  #check <- sum.aggXday %>% filter(date==run.code)
  #head(check)
  
allowed <- union(unique(run.data$actor), unique(run.data$subject))

# filter df1 to rows whose actor AND subject are in df2 somewhere
ddyad.list_filtered <- dyad.list %>%
  filter(actor %in% allowed & subject %in% allowed)
 

bird.list_filtered <- bird.list[bird.list %in% allowed & !is.na(bird.list)]


  data_alldyads <- merge(ddyad.list_filtered, run.data, all.x=TRUE,
                       by=c("actor", "subject")) 
  
  #head(data_alldyads)
  #length(unique(data_alldyads$dyadID)) # with 21 birds should be 420 and with 22 462 dyads
  
  data_alldyads$date<-run.code
  
 #fill newly-merged data with 0's where no interactions
 data_alldyads[is.na(data_alldyads)] <- 0
  
   #print a check
  check <- length(data_alldyads$actor)
  print(check)

# matrix  
  mx <- reshape2::dcast(data_alldyads, actor~subject, value.var="n.aggXbin") #head(ref.behavior1.mx)
  mx[is.na(mx)] <- 0 #for linearity measure, matrix needs to be fully filled, no NAs
  mx <- matrix.please(mx)
  
  ## dominance
dom_ec <- dom_ec(mx)

## Betweenness centrality
# network
  graph <- graph_from_adjacency_matrix(mx, "directed", weighted=T, diag=F)

# betweenness centrality
between <- betweenness(graph, 
                          directed = TRUE, 
                          weights = E(graph)$weight)



# combine in dataframe
pool <- cbind.data.frame(run.code, 
                         bird.list_filtered,
                         dom_ec,
                         between)

  ec_rank<-rbind(ec_rank, pool )

}

end.time <- Sys.time()

#Time to run:
end.time - start.time

# check 
head(ec_rank)
length(unique(ec_rank$run.code))
check<-ec_rank %>% group_by(run.code) %>% summarize(length(bird.list_filtered))

ec_rankXbin<-ec_rank %>%
  rename(bin=run.code, id=bird.list_filtered)

# higher rank is higher position in the hierarchy
head(ec_rankXbin)


ec_rankXbin <- arrange(ec_rankXbin, bin, dom_ec)

MOPA_bin_rank_2022G2 <- ec_rankXbin %>%
  group_by(bin) %>%
  mutate(rank = dense_rank(dom_ec)) %>%
  ungroup()


#write.csv(MOPA_bin_rank_2022G2,  "MOPA_bin_rank_2022G2.csv", row.names = FALSE)


```
