---
title: "Supp mat observed motif count compared to random plots"
author: 
date: 
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#load packages
```{r}
library(readr)
library(dplyr)
library(tidyr)
library(here)
library(patchwork)

```


#rule spec 75%
##Combine individual randomized and observed dfs respectively
```{r}
#IMPORTANT
#Only run chunk if not using provided dfs (tdr_count_rand_comb_rule_spec_75_perc.csv and tdr_count_real_comb_rule_spec_75_perc.csv) and want to combine each randomized dataframe and each observed dataframe individually. If so be sure to run each rule's script up to "Calculate Observed Rule"





#2021
#retal
masterframe_retaliation.df_2021 <- read_csv("retal2021countmaster_rule_spec_75perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_retaliation.df_2021$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_retaliation_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_retaliations <- length(which(masterframe_retaliation.df_2021$rule_followed == "yes" & 
                                        masterframe_retaliation.df_2021$runIDs == j))
random_times_retaliation_followed <- c(random_times_retaliation_followed, len_of_retaliations)
}
#turn vector into a dataframe for analysis
rand_retal_.frame <- data.frame(random_times_retaliation_followed)


#pile
masterframe_pileon.df_2021 <- read_csv("pileon2021countmaster_rule_spec_75perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_pileon.df_2021$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_pileon_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_pileons <- length(which(masterframe_pileon.df_2021$rule_followed == "yes" & 
                                        masterframe_pileon.df_2021$runIDs == j))
random_times_pileon_followed <- c(random_times_pileon_followed, len_of_pileons)
}
#turn vector into a dataframe for analysis
rand_pile_.frame <- data.frame(random_times_pileon_followed)


#pass
masterframe_passalong.df_2021 <- read_csv("passalong2021countmaster_rule_spec_75perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_passalong.df_2021$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_passalong_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_passalongs <- length(which(masterframe_passalong.df_2021$rule_followed == "yes" & 
                                        masterframe_passalong.df_2021$runIDs == j))
random_times_passalong_followed <- c(random_times_passalong_followed, len_of_passalongs)
}
#turn vector into a dataframe for analysis
rand_pass_.frame <- data.frame(random_times_passalong_followed)




#opp
masterframe_opp.df_2021 <- read_csv("opp2021countmaster_rule_spec_75perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_opp.df_2021$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_opp_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_opps <- length(which(masterframe_opp.df_2021$rule_followed == "yes" & 
                                        masterframe_opp.df_2021$runIDs == j))
random_times_opp_followed <- c(random_times_opp_followed, len_of_opps)
}
#turn vector into a dataframe for analysis
rand_opp_.frame <- data.frame(random_times_opp_followed)


#pun
masterframe_pun.df_2021 <- read_csv("punishment2021countmaster_rule_spec_75perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_pun.df_2021$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_pun_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_puns <- length(which(masterframe_pun.df_2021$rule_followed == "yes" & 
                                        masterframe_pun.df_2021$runIDs == j))
random_times_pun_followed <- c(random_times_pun_followed, len_of_puns)
}
#turn vector into a dataframe for analysis
rand_pun_.frame <- data.frame(random_times_pun_followed)






#2022G1
masterframe_retaliation.df_2022G1 <- read_csv("retal2022G1countmaster_rule_spec_75perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_retaliation.df_2022G1$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_retaliation_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_retaliations <- length(which(masterframe_retaliation.df_2022G1$rule_followed == "yes" & 
                                        masterframe_retaliation.df_2022G1$runIDs == j))
random_times_retaliation_followed <- c(random_times_retaliation_followed, len_of_retaliations)
}
#turn vector into a dataframe for analysis
rand_retal_.frame2022G1 <- data.frame(random_times_retaliation_followed)




#pile
masterframe_pileon.df_2022G1 <- read_csv("pileon2022G1countmaster_rule_spec_75perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_pileon.df_2022G1$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_pileon_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_pileons <- length(which(masterframe_pileon.df_2022G1$rule_followed == "yes" & 
                                        masterframe_pileon.df_2022G1$runIDs == j))
random_times_pileon_followed <- c(random_times_pileon_followed, len_of_pileons)
}
#turn vector into a dataframe for analysis
rand_pile_.frame_2022G1 <- data.frame(random_times_pileon_followed)


#pass
masterframe_passalong.df_2022G1 <- read_csv("passalong2022G1countmaster_rule_spec_75perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_passalong.df_2022G1$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_passalong_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_passalongs <- length(which(masterframe_passalong.df_2022G1$rule_followed == "yes" & 
                                        masterframe_passalong.df_2022G1$runIDs == j))
random_times_passalong_followed <- c(random_times_passalong_followed, len_of_passalongs)
}
#turn vector into a dataframe for analysis
rand_pass_.frame_2022G1 <- data.frame(random_times_passalong_followed)




#opp
masterframe_opp.df_2022G1 <- read_csv("opp2022G1countmaster_rule_spec_75perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_opp.df_2022G1$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_opp_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_opps <- length(which(masterframe_opp.df_2022G1$rule_followed == "yes" & 
                                        masterframe_opp.df_2022G1$runIDs == j))
random_times_opp_followed <- c(random_times_opp_followed, len_of_opps)
}
#turn vector into a dataframe for analysis
rand_opp_.frame_2022G1 <- data.frame(random_times_opp_followed)


#pun
masterframe_pun.df_2022G1 <- read_csv("punishment2022G1countmaster_rule_spec_75perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_pun.df_2022G1$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_pun_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_puns <- length(which(masterframe_pun.df_2022G1$rule_followed == "yes" & 
                                        masterframe_pun.df_2022G1$runIDs == j))
random_times_pun_followed <- c(random_times_pun_followed, len_of_puns)
}
#turn vector into a dataframe for analysis
rand_pun_.frame_2022G1 <- data.frame(random_times_pun_followed)









#2022G2

#2022G2
masterframe_retaliation.df_2022G2 <- read_csv("retal2022G2countmaster_rule_spec_75perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_retaliation.df_2022G2$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_retaliation_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_retaliations <- length(which(masterframe_retaliation.df_2022G2$rule_followed == "yes" & 
                                        masterframe_retaliation.df_2022G2$runIDs == j))
random_times_retaliation_followed <- c(random_times_retaliation_followed, len_of_retaliations)
}
#turn vector into a dataframe for analysis
rand_retal_.frame2022G2 <- data.frame(random_times_retaliation_followed)




#pile
masterframe_pileon.df_2022G2 <- read_csv("pileon2022G2countmaster_rule_spec_75_perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_pileon.df_2022G2$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_pileon_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_pileons <- length(which(masterframe_pileon.df_2022G2$rule_followed == "yes" & 
                                        masterframe_pileon.df_2022G2$runIDs == j))
random_times_pileon_followed <- c(random_times_pileon_followed, len_of_pileons)
}
#turn vector into a dataframe for analysis
rand_pile_.frame_2022G2 <- data.frame(random_times_pileon_followed)


#pass
masterframe_passalong.df_2022G2 <- read_csv("passalong2022G2countmaster_rule_spec_75perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_passalong.df_2022G2$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_passalong_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_passalongs <- length(which(masterframe_passalong.df_2022G2$rule_followed == "yes" & 
                                        masterframe_passalong.df_2022G2$runIDs == j))
random_times_passalong_followed <- c(random_times_passalong_followed, len_of_passalongs)
}
#turn vector into a dataframe for analysis
rand_pass_.frame_2022G2 <- data.frame(random_times_passalong_followed)




#opp
masterframe_opp.df_2022G2 <- read_csv("opp2022G2countmaster_rule_spec_75perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_opp.df_2022G2$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_opp_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_opps <- length(which(masterframe_opp.df_2022G2$rule_followed == "yes" & 
                                        masterframe_opp.df_2022G2$runIDs == j))
random_times_opp_followed <- c(random_times_opp_followed, len_of_opps)
}
#turn vector into a dataframe for analysis
rand_opp_.frame_2022G2 <- data.frame(random_times_opp_followed)


#pun
masterframe_pun.df_2022G2 <- read_csv("punishment2022G2countmaster_rule_spec_75perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_pun.df_2022G2$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_pun_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_puns <- length(which(masterframe_pun.df_2022G2$rule_followed == "yes" & 
                                        masterframe_pun.df_2022G2$runIDs == j))
random_times_pun_followed <- c(random_times_pun_followed, len_of_puns)
}
#turn vector into a dataframe for analysis
rand_pun_.frame_2022G2 <- data.frame(random_times_pun_followed)








#make all column names consistent
colnames(rand_retal_.frame)[1] <- 'count_of_rule_follows'
colnames(rand_retal_.frame2022G1)[1] <- 'count_of_rule_follows'
colnames(rand_retal_.frame2022G2)[1] <- 'count_of_rule_follows'
colnames(rand_pile_.frame)[1] <- 'count_of_rule_follows'
colnames(rand_pile_.frame_2022G1)[1] <- 'count_of_rule_follows'
colnames(rand_pile_.frame_2022G2)[1] <- 'count_of_rule_follows'
colnames(rand_pass_.frame)[1] <- 'count_of_rule_follows'
colnames(rand_pass_.frame_2022G1)[1] <- 'count_of_rule_follows'
colnames(rand_pass_.frame_2022G2)[1] <- 'count_of_rule_follows'
colnames(rand_opp_.frame)[1] <- 'count_of_rule_follows'
colnames(rand_opp_.frame_2022G1)[1] <- 'count_of_rule_follows'
colnames(rand_opp_.frame_2022G2)[1] <- 'count_of_rule_follows'
colnames(rand_pun_.frame)[1] <- 'count_of_rule_follows'
colnames(rand_pun_.frame_2022G1)[1] <- 'count_of_rule_follows'
colnames(rand_pun_.frame_2022G2)[1] <- 'count_of_rule_follows'




#bind all dfs containing the counts of motif execution from the randomized reference models. bind possible grouping factors
tdr_count_rand_comb <- rbind(
  cbind(rand_retal_.frame, Group = "Retalialtion Group 1", groupid = "Group 1", rule = "Retaliation"),
  cbind(rand_retal_.frame2022G1, Group = "Retaliation Group 2", groupid = "Group 2", rule = "Retaliation"),
  cbind(rand_retal_.frame2022G2, Group = "Retaliation Group 3", groupid = "Group 3", rule = "Retaliation"),
  cbind(rand_pile_.frame, Group = "Pile-on Group 1", groupid = "Group 1", rule = "Pile-on"),  
  cbind(rand_pile_.frame_2022G1, Group = "Pile-on Group 2", groupid = "Group 2", rule = "Pile-on"),
  cbind(rand_pile_.frame_2022G2, Group = "Pile-on Group 3", groupid = "Group 3", rule = "Pile-on"),
  cbind(rand_pass_.frame, Group = "Pass-along Group 1", groupid = "Group 1", rule = "Pass-along"),
  cbind(rand_pass_.frame_2022G1, Group = "Pass-along Group 2", groupid = "Group 2", rule = "Pass-along"),
  cbind(rand_pass_.frame_2022G2, Group = "Pass-along Group 3", groupid = "Group 3", rule = "Pass-along"),
  cbind(rand_opp_.frame, Group = "Opportunism Group 1", groupid = "Group 1", rule = "Opportunism"),
  cbind(rand_opp_.frame_2022G1, Group = "Opportunism Group 2", groupid = "Group 2", rule = "Opportunism"),
  cbind(rand_opp_.frame_2022G2, Group = "Opportunism Group 3", groupid = "Group 3", rule = "Opportunism"),
  cbind(rand_pun_.frame, Group = "Punishment Group 1", groupid = "Group 1", rule = "Punishment"),
  cbind(rand_pun_.frame_2022G1, Group = "Punishment Group 2", groupid = "Group 2", rule = "Punishment"),
  cbind(rand_pun_.frame_2022G2, Group = "Punishment Group 3", groupid = "Group 3", rule = "Punishment")

)




#write.csv(tdr_count_rand_comb, "tdr_count_rand_comb_rule_spec_75_perc.csv", row.names = FALSE)





#make all column names consistent
colnames(retal_real.frame)[1] <- 'count_of_rule_follows'
colnames(retal_real.frame2022G1)[1] <- 'count_of_rule_follows'
colnames(retal_real.frame2022G2)[1] <- 'count_of_rule_follows'
colnames(pile_real.frame_2021)[1] <- 'count_of_rule_follows'
colnames(pile_real.frame_2022G1)[1] <- 'count_of_rule_follows'
colnames(pile_real.frame_2022G2)[1] <- 'count_of_rule_follows'
colnames(pass_along_real_frame_2021)[1] <- 'count_of_rule_follows'
colnames(pass_along_real_2022G1_frame)[1] <- 'count_of_rule_follows'
colnames(pass_along_real_2022G2_frame)[1] <- 'count_of_rule_follows'
colnames(opp_real_2021_frame)[1] <- 'count_of_rule_follows'
colnames(opp_real_2022G1_frame)[1] <- 'count_of_rule_follows'
colnames(opp_real_2022G2_frame)[1] <- 'count_of_rule_follows'
colnames(punishment_real_2021_frame)[1] <- 'count_of_rule_follows'
colnames(punishment_real_2022G1_frame)[1] <- 'count_of_rule_follows'
colnames(punishment_real_2022G2_frame)[1] <- 'count_of_rule_follows'

# #bind all dfs containing the counts of motif execution from observed data. bind possible grouping factors
tdr_count_real_comb<- rbind(
  cbind(retal_real.frame, Group = "Retalialtion Group 1", groupid = "Group 1", rule = "Retaliation"),
  cbind(retal_real.frame2022G1, Group = "Retaliation Group 2", groupid = "Group 2", rule = "Retaliation"),
  cbind(retal_real.frame2022G2, Group = "Retaliation Group 3", groupid = "Group 3", rule = "Retaliation"),
  cbind(pile_real.frame_2021, Group = "Pile-on Group 1", groupid = "Group 1", rule = "Pile-on"),  
  cbind(pile_real.frame_2022G1, Group = "Pile-on Group 2", groupid = "Group 2", rule = "Pile-on"),
  cbind(pile_real.frame_2022G2, Group = "Pile-on Group 3", groupid = "Group 3", rule = "Pile-on"),
  cbind(pass_along_real_frame_2021, Group = "Pass-along Group 1", groupid = "Group 1", rule = "Pass-along"),
  cbind(pass_along_real_2022G1_frame, Group = "Pass-along Group 2", groupid = "Group 2", rule = "Pass-along"),
  cbind(pass_along_real_2022G2_frame, Group = "Pass-along Group 3", groupid = "Group 3", rule = "Pass-along"),
  cbind(opp_real_2021_frame, Group = "Opportunism Group 1", groupid = "Group 1", rule = "Opportunism"),
  cbind(opp_real_2022G1_frame, Group = "Opportunism Group 2", groupid = "Group 2", rule = "Opportunism"),
  cbind(opp_real_2022G2_frame, Group = "Opportunism Group 3", groupid = "Group 3", rule = "Opportunism"),
  cbind(punishment_real_2021_frame, Group = "Punishment Group 1", groupid = "Group 1", rule = "Punishment"),
  cbind(punishment_real_2022G1_frame, Group = "Punishment Group 2", groupid = "Group 2", rule = "Punishment"),
  cbind(punishment_real_2022G2_frame, Group = "Punishment Group 3", groupid = "Group 3", rule = "Punishment")

)

#write.csv(tdr_count_real_comb,  "tdr_count_real_comb_rule_spec_75_perc.csv", row.names = FALSE)



```

##Plot observed motif count compared to random violin and barplots
```{r}
#Load randomized df
tdr_count_rand_comb <- read_csv("tdr_count_rand_comb_rule_spec_75_perc.csv")


#calculate means of randomized motif execution for each rule
tdr_count_rand_comb_mean <- tdr_count_rand_comb  %>% group_by(groupid, rule) %>% summarise(mean_value = mean(count_of_rule_follows, na.rm = TRUE))


#Load observed df
tdr_count_real_comb <- read_csv("tdr_count_real_comb_rule_spec_75_perc.csv")



#set order in which rules should be displayed in the plot
custom_order <- c("Retaliation","Pile-on",  "Pass-along", "Opportunism", "Punishment")
tdr_count_rand_comb$rule <- factor(tdr_count_rand_comb$rule, levels = custom_order)

tdr_count_real_comb$rule <- factor(tdr_count_real_comb$rule, levels = custom_order)

#compressed violin version

rule_order <- c("Retaliation","Pile-on","Pass-along","Opportunism","Punishment")
# Use only rules that actually exist in data just in case but keep the specified order
present_rules <- intersect(rule_order, unique(tdr_count_rand_comb$rule))
groups <- sort(unique(tdr_count_rand_comb$groupid))

#Build axis order with rules clustered together and then g1 g2 g3 of each 
#makes it so goes through groups first then rules
#looks like Retaliation – G1,G2,G3 | Pile-on – G1,G2,G3 | ...
axis_order <- expand.grid(
  groupid = groups,
  rule    = present_rules,
  KEEP.OUT.ATTRS = FALSE
) |>
  mutate(xcat = paste0(rule, "–", groupid)) |>
  pull(xcat)
#make sure only pulling out appropriate rules and groups

tdr_rand2 <- tdr_count_rand_comb |>
  filter(rule %in% present_rules, groupid %in% groups) |>
  mutate(
    rule = factor(rule, levels = present_rules),
    xcat = factor(paste0(rule, "–", groupid), levels = axis_order)
  )
#same thing for observed motif execution
tdr_real2 <- tdr_count_real_comb |>
  filter(rule %in% present_rules, groupid %in% groups) |>
  mutate(
    rule = factor(rule, levels = present_rules),
    xcat = factor(paste0(rule, "–", groupid), levels = axis_order)
  )

#create separator positions between rule blocks
g_per_rule    <- length(groups)
sep_positions <- (seq_len(length(present_rules) - 1) * g_per_rule) + 0.5

rule_violin <- ggplot(tdr_rand2, aes(x = xcat, y = count_of_rule_follows)) +
#separators behind violins
geom_vline(xintercept = sep_positions, linetype = "dashed",
             color = "grey60", linewidth = 0.8) +
#randomized reference as gray violins
geom_violin(fill = "grey70", color = "black", alpha = 1,
              width = 0.9, trim = FALSE, scale = "width") +
#mean bar across each violin
   stat_summary(fun = mean, geom = "crossbar", width = 0.6,
              color = "black", linewidth = 0.3)+
#observed value as a red horizontal line across each category
  geom_segment(
    data = tdr_real2,
    aes(x = as.numeric(xcat) - 0.35,
        xend = as.numeric(xcat) + 0.35,
        y = count_of_rule_follows,
        yend = count_of_rule_follows),
    inherit.aes = FALSE,
    linewidth = 1.2, color = "red"
  ) +
  labs(
    x = NULL,
    y = "Count of motif execution", 
    title = "Observed vs. randomized reference by rule and group (violin)"
  ) +
  theme_minimal(base_size = 20) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.8),
    axis.text.x  = element_text(angle = 45, hjust = 1),
    axis.ticks.x = element_line()
  )

rule_violin
#15 by 12 







#short bar plot to attach
head(tdr_count_rand_comb_mean)
head(tdr_count_real_comb)


tdr_count_both_comb <- tdr_count_rand_comb_mean %>%
  left_join(tdr_count_real_comb, by = c("groupid", "rule"))
#obtain the difference between observed count and the mean of the reference distribution
tdr_count_both_comb$dif <- tdr_count_both_comb$count_of_rule_follows - tdr_count_both_comb$mean_value



rule_order <- c("Retaliation","Pile-on","Pass-along","Opportunism","Punishment")
# Use only rules that actually exist in data just in case but keep the specified order
present_rules <- intersect(rule_order, unique(tdr_count_both_comb$rule))
groups <- sort(unique(tdr_count_both_comb$groupid))

#Build axis order with rules clustered together and then g1 g2 g3 of each 
#makes it so goes through groups first then rules
#looks like Retaliation – G1,G2,G3 | Pile-on – G1,G2,G3 | ...
axis_order <- expand.grid(
  groupid = groups,
  rule    = present_rules,
  KEEP.OUT.ATTRS = FALSE
) |>
  mutate(xcat = paste0(rule, "–", groupid)) |>
  pull(xcat)
#make sure only pulling out appropriate rules and groups

tdr_count_both_comb2 <- tdr_count_both_comb |>
  filter(rule %in% present_rules, groupid %in% groups) |>
  mutate(
    rule = factor(rule, levels = present_rules),
    xcat = factor(paste0(rule, "–", groupid), levels = axis_order)
  )

#create separator positions between rule blocks
g_per_rule    <- length(groups)
sep_positions <- (seq_len(length(present_rules) - 1) * g_per_rule) + 0.5



rule_bar <- ggplot(tdr_count_both_comb2, aes(x = xcat, y = dif)) +
  #separators behind bars
  geom_vline(xintercept = sep_positions, linetype = "dashed",
             color = "grey60", linewidth = 0.8) +
  #bars from supplied y
  geom_col(fill = "grey60", color = "black") +
  labs(
    x = NULL,
    y = "Count of motif execution",
    title = "Observed vs. randomized reference by rule and group (bar)"
  ) +
  theme_minimal(base_size = 20) +
  scale_y_continuous(limits = c(-450, 100)) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.8),
    axis.text.x  = element_text(angle = 90, hjust = 1),
    axis.ticks.x = element_line()
  )

rule_bar





#keep the top plot as is visually, just hide its x label ticks so they don't repeat
rule_violin_top <- rule_violin +
  theme(
    axis.title.x = element_blank(),
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    plot.margin  = margin(5.5, 5.5, 0, 5.5)  
  )

#keep bar plot exactly as is but tweak bottom margin
rule_bar_bottom <- rule_bar +
  theme(
    plot.margin = margin(0, 5.5, 5.5, 5.5)
  )

#stack two plots together. The violins on top (bigger), bars below (smaller)
combined <- rule_violin_top / rule_bar_bottom +
  plot_layout(heights = c(2, 1), guides = "keep")

combined
#12 by 10

#P values added manually in inkscape for easier formatting

```










#global 95%
##Combine individual randomized and observed dfs respectively
```{r}

#IMPORTANT
#Only run chunk if not using provided dfs (tdr_count_rand_comb_global_95_perc.csv and tdr_count_real_comb_global_95_perc.csv) and want to combine each randomized dataframe and each observed dataframe individually. If so be sure to run each rule's script up to "Calculate Observed Rule"


#2021
#retal
masterframe_retaliation.df_2021 <- read_csv("retal2021countmaster_global_95perc.csv")
#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_retaliation.df_2021$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_retaliation_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_retaliations <- length(which(masterframe_retaliation.df_2021$rule_followed == "yes" & 
                                        masterframe_retaliation.df_2021$runIDs == j))
random_times_retaliation_followed <- c(random_times_retaliation_followed, len_of_retaliations)
}
#turn vector into a dataframe for analysis
rand_retal_.frame <- data.frame(random_times_retaliation_followed)


#pile
masterframe_pileon.df_2021 <- read_csv("pileon2021countmaster_global_95perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_pileon.df_2021$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_pileon_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_pileons <- length(which(masterframe_pileon.df_2021$rule_followed == "yes" & 
                                        masterframe_pileon.df_2021$runIDs == j))
random_times_pileon_followed <- c(random_times_pileon_followed, len_of_pileons)
}
#turn vector into a dataframe for analysis
rand_pile_.frame <- data.frame(random_times_pileon_followed)


#pass
masterframe_passalong.df_2021 <- read_csv("passalong2021countmaster_global_95perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_passalong.df_2021$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_passalong_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_passalongs <- length(which(masterframe_passalong.df_2021$rule_followed == "yes" & 
                                        masterframe_passalong.df_2021$runIDs == j))
random_times_passalong_followed <- c(random_times_passalong_followed, len_of_passalongs)
}
#turn vector into a dataframe for analysis
rand_pass_.frame <- data.frame(random_times_passalong_followed)




#opp
masterframe_opp.df_2021 <- read_csv("opp2021countmaster_global_95perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_opp.df_2021$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_opp_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_opps <- length(which(masterframe_opp.df_2021$rule_followed == "yes" & 
                                        masterframe_opp.df_2021$runIDs == j))
random_times_opp_followed <- c(random_times_opp_followed, len_of_opps)
}
#turn vector into a dataframe for analysis
rand_opp_.frame <- data.frame(random_times_opp_followed)


#pun
masterframe_pun.df_2021 <- read_csv("punishment2021countmaster_global_95perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_pun.df_2021$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_pun_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_puns <- length(which(masterframe_pun.df_2021$rule_followed == "yes" & 
                                        masterframe_pun.df_2021$runIDs == j))
random_times_pun_followed <- c(random_times_pun_followed, len_of_puns)
}
#turn vector into a dataframe for analysis
rand_pun_.frame <- data.frame(random_times_pun_followed)






#2022G1
masterframe_retaliation.df_2022G1 <- read_csv("retal2022G1countmaster_global_95perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_retaliation.df_2022G1$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_retaliation_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_retaliations <- length(which(masterframe_retaliation.df_2022G1$rule_followed == "yes" & 
                                        masterframe_retaliation.df_2022G1$runIDs == j))
random_times_retaliation_followed <- c(random_times_retaliation_followed, len_of_retaliations)
}
#turn vector into a dataframe for analysis
rand_retal_.frame2022G1 <- data.frame(random_times_retaliation_followed)




#pile
masterframe_pileon.df_2022G1 <- read_csv("pileon2022G1countmaster_global_95perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_pileon.df_2022G1$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_pileon_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_pileons <- length(which(masterframe_pileon.df_2022G1$rule_followed == "yes" & 
                                        masterframe_pileon.df_2022G1$runIDs == j))
random_times_pileon_followed <- c(random_times_pileon_followed, len_of_pileons)
}
#turn vector into a dataframe for analysis
rand_pile_.frame_2022G1 <- data.frame(random_times_pileon_followed)


#pass
masterframe_passalong.df_2022G1 <- read_csv("passalong2022G1countmaster_global_95perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_passalong.df_2022G1$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_passalong_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_passalongs <- length(which(masterframe_passalong.df_2022G1$rule_followed == "yes" & 
                                        masterframe_passalong.df_2022G1$runIDs == j))
random_times_passalong_followed <- c(random_times_passalong_followed, len_of_passalongs)
}
#turn vector into a dataframe for analysis
rand_pass_.frame_2022G1 <- data.frame(random_times_passalong_followed)




#opp
masterframe_opp.df_2022G1 <- read_csv("opp2022G1countmaster_global_95perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_opp.df_2022G1$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_opp_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_opps <- length(which(masterframe_opp.df_2022G1$rule_followed == "yes" & 
                                        masterframe_opp.df_2022G1$runIDs == j))
random_times_opp_followed <- c(random_times_opp_followed, len_of_opps)
}
#turn vector into a dataframe for analysis
rand_opp_.frame_2022G1 <- data.frame(random_times_opp_followed)


#pun
masterframe_pun.df_2022G1 <- read_csv("punishment2022G1countmaster_global_95perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_pun.df_2022G1$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_pun_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_puns <- length(which(masterframe_pun.df_2022G1$rule_followed == "yes" & 
                                        masterframe_pun.df_2022G1$runIDs == j))
random_times_pun_followed <- c(random_times_pun_followed, len_of_puns)
}
#turn vector into a dataframe for analysis
rand_pun_.frame_2022G1 <- data.frame(random_times_pun_followed)









#2022G2

#2022G2
masterframe_retaliation.df_2022G2 <- read_csv("retal2022G2countmaster_global_95perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_retaliation.df_2022G2$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_retaliation_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_retaliations <- length(which(masterframe_retaliation.df_2022G2$rule_followed == "yes" & 
                                        masterframe_retaliation.df_2022G2$runIDs == j))
random_times_retaliation_followed <- c(random_times_retaliation_followed, len_of_retaliations)
}
#turn vector into a dataframe for analysis
rand_retal_.frame2022G2 <- data.frame(random_times_retaliation_followed)




#pile
masterframe_pileon.df_2022G2 <- read_csv("pileon2022G2countmaster_global_95perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_pileon.df_2022G2$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_pileon_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_pileons <- length(which(masterframe_pileon.df_2022G2$rule_followed == "yes" & 
                                        masterframe_pileon.df_2022G2$runIDs == j))
random_times_pileon_followed <- c(random_times_pileon_followed, len_of_pileons)
}
#turn vector into a dataframe for analysis
rand_pile_.frame_2022G2 <- data.frame(random_times_pileon_followed)


#pass
masterframe_passalong.df_2022G2 <- read_csv("passalong2022G2countmaster_global_95perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_passalong.df_2022G2$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_passalong_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_passalongs <- length(which(masterframe_passalong.df_2022G2$rule_followed == "yes" & 
                                        masterframe_passalong.df_2022G2$runIDs == j))
random_times_passalong_followed <- c(random_times_passalong_followed, len_of_passalongs)
}
#turn vector into a dataframe for analysis
rand_pass_.frame_2022G2 <- data.frame(random_times_passalong_followed)




#opp
masterframe_opp.df_2022G2 <- read_csv("opp2022G2countmaster_global_95perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_opp.df_2022G2$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_opp_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_opps <- length(which(masterframe_opp.df_2022G2$rule_followed == "yes" & 
                                        masterframe_opp.df_2022G2$runIDs == j))
random_times_opp_followed <- c(random_times_opp_followed, len_of_opps)
}
#turn vector into a dataframe for analysis
rand_opp_.frame_2022G2 <- data.frame(random_times_opp_followed)


#pun
masterframe_pun.df_2022G2 <- read_csv("punishment2022G2countmaster_global_95perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_pun.df_2022G2$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_pun_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_puns <- length(which(masterframe_pun.df_2022G2$rule_followed == "yes" & 
                                        masterframe_pun.df_2022G2$runIDs == j))
random_times_pun_followed <- c(random_times_pun_followed, len_of_puns)
}
#turn vector into a dataframe for analysis
rand_pun_.frame_2022G2 <- data.frame(random_times_pun_followed)








#make all column names consistent
colnames(rand_retal_.frame)[1] <- 'count_of_rule_follows'
colnames(rand_retal_.frame2022G1)[1] <- 'count_of_rule_follows'
colnames(rand_retal_.frame2022G2)[1] <- 'count_of_rule_follows'
colnames(rand_pile_.frame)[1] <- 'count_of_rule_follows'
colnames(rand_pile_.frame_2022G1)[1] <- 'count_of_rule_follows'
colnames(rand_pile_.frame_2022G2)[1] <- 'count_of_rule_follows'
colnames(rand_pass_.frame)[1] <- 'count_of_rule_follows'
colnames(rand_pass_.frame_2022G1)[1] <- 'count_of_rule_follows'
colnames(rand_pass_.frame_2022G2)[1] <- 'count_of_rule_follows'
colnames(rand_opp_.frame)[1] <- 'count_of_rule_follows'
colnames(rand_opp_.frame_2022G1)[1] <- 'count_of_rule_follows'
colnames(rand_opp_.frame_2022G2)[1] <- 'count_of_rule_follows'
colnames(rand_pun_.frame)[1] <- 'count_of_rule_follows'
colnames(rand_pun_.frame_2022G1)[1] <- 'count_of_rule_follows'
colnames(rand_pun_.frame_2022G2)[1] <- 'count_of_rule_follows'






#bind all dfs containing the counts of motif execution from the randomized reference models. bind possible grouping factors
tdr_count_rand_comb <- rbind(
  cbind(rand_retal_.frame, Group = "Retalialtion Group 1", groupid = "Group 1", rule = "Retaliation"),
  cbind(rand_retal_.frame2022G1, Group = "Retaliation Group 2", groupid = "Group 2", rule = "Retaliation"),
  cbind(rand_retal_.frame2022G2, Group = "Retaliation Group 3", groupid = "Group 3", rule = "Retaliation"),
  cbind(rand_pile_.frame, Group = "Pile-on Group 1", groupid = "Group 1", rule = "Pile-on"),  
  cbind(rand_pile_.frame_2022G1, Group = "Pile-on Group 2", groupid = "Group 2", rule = "Pile-on"),
  cbind(rand_pile_.frame_2022G2, Group = "Pile-on Group 3", groupid = "Group 3", rule = "Pile-on"),
  cbind(rand_pass_.frame, Group = "Pass-along Group 1", groupid = "Group 1", rule = "Pass-along"),
  cbind(rand_pass_.frame_2022G1, Group = "Pass-along Group 2", groupid = "Group 2", rule = "Pass-along"),
  cbind(rand_pass_.frame_2022G2, Group = "Pass-along Group 3", groupid = "Group 3", rule = "Pass-along"),
  cbind(rand_opp_.frame, Group = "Opportunism Group 1", groupid = "Group 1", rule = "Opportunism"),
  cbind(rand_opp_.frame_2022G1, Group = "Opportunism Group 2", groupid = "Group 2", rule = "Opportunism"),
  cbind(rand_opp_.frame_2022G2, Group = "Opportunism Group 3", groupid = "Group 3", rule = "Opportunism"),
  cbind(rand_pun_.frame, Group = "Punishment Group 1", groupid = "Group 1", rule = "Punishment"),
  cbind(rand_pun_.frame_2022G1, Group = "Punishment Group 2", groupid = "Group 2", rule = "Punishment"),
  cbind(rand_pun_.frame_2022G2, Group = "Punishment Group 3", groupid = "Group 3", rule = "Punishment")

)

#write.csv(tdr_count_rand_comb,  "tdr_count_rand_comb_global_95_perc.csv", row.names = FALSE)




#make all column names consistent
colnames(retal_real.frame)[1] <- 'count_of_rule_follows'
colnames(retal_real.frame2022G1)[1] <- 'count_of_rule_follows'
colnames(retal_real.frame2022G2)[1] <- 'count_of_rule_follows'
colnames(pile_real.frame_2021)[1] <- 'count_of_rule_follows'
colnames(pile_real.frame_2022G1)[1] <- 'count_of_rule_follows'
colnames(pile_real.frame_2022G2)[1] <- 'count_of_rule_follows'
colnames(pass_along_real_frame_2021)[1] <- 'count_of_rule_follows'
colnames(pass_along_real_2022G1_frame)[1] <- 'count_of_rule_follows'
colnames(pass_along_real_2022G2_frame)[1] <- 'count_of_rule_follows'
colnames(opp_real_2021_frame)[1] <- 'count_of_rule_follows'
colnames(opp_real_2022G1_frame)[1] <- 'count_of_rule_follows'
colnames(opp_real_2022G2_frame)[1] <- 'count_of_rule_follows'
colnames(punishment_real_2021_frame)[1] <- 'count_of_rule_follows'
colnames(punishment_real_2022G1_frame)[1] <- 'count_of_rule_follows'
colnames(punishment_real_2022G2_frame)[1] <- 'count_of_rule_follows'

# #bind all dfs containing the counts of motif execution from observed data. bind possible grouping factors
tdr_count_real_comb<- rbind(
  cbind(retal_real.frame, Group = "Retalialtion Group 1", groupid = "Group 1", rule = "Retaliation"),
  cbind(retal_real.frame2022G1, Group = "Retaliation Group 2", groupid = "Group 2", rule = "Retaliation"),
  cbind(retal_real.frame2022G2, Group = "Retaliation Group 3", groupid = "Group 3", rule = "Retaliation"),
  cbind(pile_real.frame_2021, Group = "Pile-on Group 1", groupid = "Group 1", rule = "Pile-on"),  
  cbind(pile_real.frame_2022G1, Group = "Pile-on Group 2", groupid = "Group 2", rule = "Pile-on"),
  cbind(pile_real.frame_2022G2, Group = "Pile-on Group 3", groupid = "Group 3", rule = "Pile-on"),
  cbind(pass_along_real_frame_2021, Group = "Pass-along Group 1", groupid = "Group 1", rule = "Pass-along"),
  cbind(pass_along_real_2022G1_frame, Group = "Pass-along Group 2", groupid = "Group 2", rule = "Pass-along"),
  cbind(pass_along_real_2022G2_frame, Group = "Pass-along Group 3", groupid = "Group 3", rule = "Pass-along"),
  cbind(opp_real_2021_frame, Group = "Opportunism Group 1", groupid = "Group 1", rule = "Opportunism"),
  cbind(opp_real_2022G1_frame, Group = "Opportunism Group 2", groupid = "Group 2", rule = "Opportunism"),
  cbind(opp_real_2022G2_frame, Group = "Opportunism Group 3", groupid = "Group 3", rule = "Opportunism"),
  cbind(punishment_real_2021_frame, Group = "Punishment Group 1", groupid = "Group 1", rule = "Punishment"),
  cbind(punishment_real_2022G1_frame, Group = "Punishment Group 2", groupid = "Group 2", rule = "Punishment"),
  cbind(punishment_real_2022G2_frame, Group = "Punishment Group 3", groupid = "Group 3", rule = "Punishment")

)

#write.csv(tdr_count_real_comb,  "tdr_count_real_comb_global_95_perc.csv", row.names = FALSE)




```

##Plot observed motif count compared to random violin and barplots
```{r}

#Load randomized df
tdr_count_rand_comb <- read_csv("tdr_count_rand_comb_global_95_perc.csv")


#calculate means of randomized motif execution for each rule
tdr_count_rand_comb_mean <- tdr_count_rand_comb  %>% group_by(groupid, rule) %>% summarise(mean_value = mean(count_of_rule_follows, na.rm = TRUE))




#Load observed df
tdr_count_real_comb <- read_csv("tdr_count_real_comb_global_95_perc.csv")



#set order in which rules should be displayed in the plot
custom_order <- c("Retaliation","Pile-on",  "Pass-along", "Opportunism", "Punishment")
tdr_count_rand_comb$rule <- factor(tdr_count_rand_comb$rule, levels = custom_order)

tdr_count_real_comb$rule <- factor(tdr_count_real_comb$rule, levels = custom_order)

#compressed violin version

rule_order <- c("Retaliation","Pile-on","Pass-along","Opportunism","Punishment")
# Use only rules that actually exist in data just in case but keep the specified order
present_rules <- intersect(rule_order, unique(tdr_count_rand_comb$rule))
groups <- sort(unique(tdr_count_rand_comb$groupid))

#Build axis order with rules clustered together and then g1 g2 g3 of each 
#makes it so goes through groups first then rules
#looks like Retaliation – G1,G2,G3 | Pile-on – G1,G2,G3 | ...
axis_order <- expand.grid(
  groupid = groups,
  rule    = present_rules,
  KEEP.OUT.ATTRS = FALSE
) |>
  mutate(xcat = paste0(rule, "–", groupid)) |>
  pull(xcat)
#make sure only pulling out appropriate rules and groups

tdr_rand2 <- tdr_count_rand_comb |>
  filter(rule %in% present_rules, groupid %in% groups) |>
  mutate(
    rule = factor(rule, levels = present_rules),
    xcat = factor(paste0(rule, "–", groupid), levels = axis_order)
  )
#same thing for observed motif execution
tdr_real2 <- tdr_count_real_comb |>
  filter(rule %in% present_rules, groupid %in% groups) |>
  mutate(
    rule = factor(rule, levels = present_rules),
    xcat = factor(paste0(rule, "–", groupid), levels = axis_order)
  )

#create separator positions between rule blocks
g_per_rule    <- length(groups)
sep_positions <- (seq_len(length(present_rules) - 1) * g_per_rule) + 0.5

rule_violin <- ggplot(tdr_rand2, aes(x = xcat, y = count_of_rule_follows)) +
#separators behind violins
geom_vline(xintercept = sep_positions, linetype = "dashed",
             color = "grey60", linewidth = 0.8) +
#randomized reference as gray violins
geom_violin(fill = "grey70", color = "black", alpha = 1,
              width = 0.9, trim = FALSE, scale = "width") +
#mean bar across each violin
   stat_summary(fun = mean, geom = "crossbar", width = 0.6,
              color = "black", linewidth = 0.3)+
#observed value as a red horizontal line across each category
  geom_segment(
    data = tdr_real2,
    aes(x = as.numeric(xcat) - 0.35,
        xend = as.numeric(xcat) + 0.35,
        y = count_of_rule_follows,
        yend = count_of_rule_follows),
    inherit.aes = FALSE,
    linewidth = 1.2, color = "red"
  ) +
  labs(
    x = NULL,
    y = "Count of motif execution", 
    title = "Observed vs. randomized reference by rule and group (violin)"
  ) +
  theme_minimal(base_size = 20) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.8),
    axis.text.x  = element_text(angle = 45, hjust = 1),
    axis.ticks.x = element_line()
  )

rule_violin
#15 by 12 or 12 by 15









#short bar plot to attach
head(tdr_count_rand_comb_mean)
head(tdr_count_real_comb)


tdr_count_both_comb <- tdr_count_rand_comb_mean %>%
  left_join(tdr_count_real_comb, by = c("groupid", "rule"))
#obtain the difference between observed count and the mean of the reference distribution
tdr_count_both_comb$dif <- tdr_count_both_comb$count_of_rule_follows - tdr_count_both_comb$mean_value



rule_order <- c("Retaliation","Pile-on","Pass-along","Opportunism","Punishment")
# Use only rules that actually exist in data just in case but keep the specified order
present_rules <- intersect(rule_order, unique(tdr_count_both_comb$rule))
groups <- sort(unique(tdr_count_both_comb$groupid))

#Build axis order with rules clustered together and then g1 g2 g3 of each 
#makes it so goes through groups first then rules
#looks like Retaliation – G1,G2,G3 | Pile-on – G1,G2,G3 | ...
axis_order <- expand.grid(
  groupid = groups,
  rule    = present_rules,
  KEEP.OUT.ATTRS = FALSE
) |>
  mutate(xcat = paste0(rule, "–", groupid)) |>
  pull(xcat)
#make sure only pulling out appropriate rules and groups

tdr_count_both_comb2 <- tdr_count_both_comb |>
  filter(rule %in% present_rules, groupid %in% groups) |>
  mutate(
    rule = factor(rule, levels = present_rules),
    xcat = factor(paste0(rule, "–", groupid), levels = axis_order)
  )

#create separator positions between rule blocks
g_per_rule    <- length(groups)
sep_positions <- (seq_len(length(present_rules) - 1) * g_per_rule) + 0.5



rule_bar <- ggplot(tdr_count_both_comb2, aes(x = xcat, y = dif)) +
  #separators behind bars
  geom_vline(xintercept = sep_positions, linetype = "dashed",
             color = "grey60", linewidth = 0.8) +
  #bars from supplied y
  geom_col(fill = "grey60", color = "black") +
  labs(
    x = NULL,
    y = "Count of motif execution",
    title = "Observed vs. randomized reference by rule and group (bar)"
  ) +
  theme_minimal(base_size = 20) +
  scale_y_continuous(limits = c(-450, 100)) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.8),
    axis.text.x  = element_text(angle = 90, hjust = 1),
    axis.ticks.x = element_line()
  )

rule_bar





#keep the top plot as is visually, just hide its x label ticks so they don't repeat
rule_violin_top <- rule_violin +
  theme(
    axis.title.x = element_blank(),
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    plot.margin  = margin(5.5, 5.5, 0, 5.5)  
  )

#keep bar plot exactly as is but tweak bottom margin
rule_bar_bottom <- rule_bar +
  theme(
    plot.margin = margin(0, 5.5, 5.5, 5.5)
  )

#stack two plots together. The violins on top (bigger), bars below (smaller)
combined <- rule_violin_top / rule_bar_bottom +
  plot_layout(heights = c(2, 1), guides = "keep")

combined
#12 by 10

#P values added manually in inkscape for easier formatting

```


#global 75%
##Combine individual randomized and observed dfs respectively
```{r}

#IMPORTANT
#Only run chunk if not using provided dfs (tdr_count_rand_comb_global_75_perc.csv and tdr_count_real_comb_global_75_perc.csv) and want to combine each randomized dataframe and each observed dataframe individually. If so be sure to run each rule's script up to "Calculate Observed Rule"




#2021
#retal
masterframe_retaliation.df_2021 <- read_csv("retal2021countmaster_global_75perc.csv")
#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_retaliation.df_2021$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_retaliation_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_retaliations <- length(which(masterframe_retaliation.df_2021$rule_followed == "yes" & 
                                        masterframe_retaliation.df_2021$runIDs == j))
random_times_retaliation_followed <- c(random_times_retaliation_followed, len_of_retaliations)
}
#turn vector into a dataframe for analysis
rand_retal_.frame <- data.frame(random_times_retaliation_followed)




#pile
masterframe_pileon.df_2021 <- read_csv("pileon2021countmaster_global_75perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_pileon.df_2021$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_pileon_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_pileons <- length(which(masterframe_pileon.df_2021$rule_followed == "yes" & 
                                        masterframe_pileon.df_2021$runIDs == j))
random_times_pileon_followed <- c(random_times_pileon_followed, len_of_pileons)
}
#turn vector into a dataframe for analysis
rand_pile_.frame <- data.frame(random_times_pileon_followed)


#pass
masterframe_passalong.df_2021 <- read_csv("passalong2021countmaster_global_75perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_passalong.df_2021$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_passalong_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_passalongs <- length(which(masterframe_passalong.df_2021$rule_followed == "yes" & 
                                        masterframe_passalong.df_2021$runIDs == j))
random_times_passalong_followed <- c(random_times_passalong_followed, len_of_passalongs)
}
#turn vector into a dataframe for analysis
rand_pass_.frame <- data.frame(random_times_passalong_followed)




#opp
masterframe_opp.df_2021 <- read_csv("opp2021countmaster_global_75perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_opp.df_2021$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_opp_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_opps <- length(which(masterframe_opp.df_2021$rule_followed == "yes" & 
                                        masterframe_opp.df_2021$runIDs == j))
random_times_opp_followed <- c(random_times_opp_followed, len_of_opps)
}
#turn vector into a dataframe for analysis
rand_opp_.frame <- data.frame(random_times_opp_followed)


#pun
masterframe_pun.df_2021 <- read_csv("punishment2021countmaster_global_75perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_pun.df_2021$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_pun_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_puns <- length(which(masterframe_pun.df_2021$rule_followed == "yes" & 
                                        masterframe_pun.df_2021$runIDs == j))
random_times_pun_followed <- c(random_times_pun_followed, len_of_puns)
}
#turn vector into a dataframe for analysis
rand_pun_.frame <- data.frame(random_times_pun_followed)






#2022G1
masterframe_retaliation.df_2022G1 <- read_csv("retal2022G1countmaster_global_75perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_retaliation.df_2022G1$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_retaliation_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_retaliations <- length(which(masterframe_retaliation.df_2022G1$rule_followed == "yes" & 
                                        masterframe_retaliation.df_2022G1$runIDs == j))
random_times_retaliation_followed <- c(random_times_retaliation_followed, len_of_retaliations)
}
#turn vector into a dataframe for analysis
rand_retal_.frame2022G1 <- data.frame(random_times_retaliation_followed)




#pile
masterframe_pileon.df_2022G1 <- read_csv("pileon2022G1countmaster_global_75perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_pileon.df_2022G1$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_pileon_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_pileons <- length(which(masterframe_pileon.df_2022G1$rule_followed == "yes" & 
                                        masterframe_pileon.df_2022G1$runIDs == j))
random_times_pileon_followed <- c(random_times_pileon_followed, len_of_pileons)
}
#turn vector into a dataframe for analysis
rand_pile_.frame_2022G1 <- data.frame(random_times_pileon_followed)


#pass
masterframe_passalong.df_2022G1 <- read_csv("passalong2022G1countmaster_global_75perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_passalong.df_2022G1$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_passalong_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_passalongs <- length(which(masterframe_passalong.df_2022G1$rule_followed == "yes" & 
                                        masterframe_passalong.df_2022G1$runIDs == j))
random_times_passalong_followed <- c(random_times_passalong_followed, len_of_passalongs)
}
#turn vector into a dataframe for analysis
rand_pass_.frame_2022G1 <- data.frame(random_times_passalong_followed)



#opp
masterframe_opp.df_2022G1 <- read_csv("opp2022G1countmaster_global_75perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_opp.df_2022G1$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_opp_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_opps <- length(which(masterframe_opp.df_2022G1$rule_followed == "yes" & 
                                        masterframe_opp.df_2022G1$runIDs == j))
random_times_opp_followed <- c(random_times_opp_followed, len_of_opps)
}
#turn vector into a dataframe for analysis
rand_opp_.frame_2022G1 <- data.frame(random_times_opp_followed)


#pun
masterframe_pun.df_2022G1 <- read_csv("punishment2022G1countmaster_global_75perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_pun.df_2022G1$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_pun_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_puns <- length(which(masterframe_pun.df_2022G1$rule_followed == "yes" & 
                                        masterframe_pun.df_2022G1$runIDs == j))
random_times_pun_followed <- c(random_times_pun_followed, len_of_puns)
}
#turn vector into a dataframe for analysis
rand_pun_.frame_2022G1 <- data.frame(random_times_pun_followed)









#2022G2

#2022G2
masterframe_retaliation.df_2022G2 <- read_csv("retal2022G2countmaster_global_75perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_retaliation.df_2022G2$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_retaliation_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_retaliations <- length(which(masterframe_retaliation.df_2022G2$rule_followed == "yes" & 
                                        masterframe_retaliation.df_2022G2$runIDs == j))
random_times_retaliation_followed <- c(random_times_retaliation_followed, len_of_retaliations)
}
#turn vector into a dataframe for analysis
rand_retal_.frame2022G2 <- data.frame(random_times_retaliation_followed)




#pile
masterframe_pileon.df_2022G2 <- read_csv("pileon2022G2countmaster_global_75perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_pileon.df_2022G2$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_pileon_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_pileons <- length(which(masterframe_pileon.df_2022G2$rule_followed == "yes" & 
                                        masterframe_pileon.df_2022G2$runIDs == j))
random_times_pileon_followed <- c(random_times_pileon_followed, len_of_pileons)
}
#turn vector into a dataframe for analysis
rand_pile_.frame_2022G2 <- data.frame(random_times_pileon_followed)


#pass
masterframe_passalong.df_2022G2 <- read_csv("passalong2022G2countmaster_global_75perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_passalong.df_2022G2$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_passalong_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_passalongs <- length(which(masterframe_passalong.df_2022G2$rule_followed == "yes" & 
                                        masterframe_passalong.df_2022G2$runIDs == j))
random_times_passalong_followed <- c(random_times_passalong_followed, len_of_passalongs)
}
#turn vector into a dataframe for analysis
rand_pass_.frame_2022G2 <- data.frame(random_times_passalong_followed)




#opp
masterframe_opp.df_2022G2 <- read_csv("opp2022G2countmaster_global_75perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_opp.df_2022G2$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_opp_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_opps <- length(which(masterframe_opp.df_2022G2$rule_followed == "yes" & 
                                        masterframe_opp.df_2022G2$runIDs == j))
random_times_opp_followed <- c(random_times_opp_followed, len_of_opps)
}
#turn vector into a dataframe for analysis
rand_opp_.frame_2022G2 <- data.frame(random_times_opp_followed)


#pun
masterframe_pun.df_2022G2 <- read_csv("punishment2022G2countmaster_global_75perc.csv")

#put run #s in an ascending vector
cum_run_IDS <- unique(masterframe_pun.df_2022G2$runIDs)

#create empty vector for number of time the motifs were executed to be stored
random_times_pun_followed <- c()

#put the number of times the motif was executed into the vector for analysis 
for (j in cum_run_IDS) {
  len_of_puns <- length(which(masterframe_pun.df_2022G2$rule_followed == "yes" & 
                                        masterframe_pun.df_2022G2$runIDs == j))
random_times_pun_followed <- c(random_times_pun_followed, len_of_puns)
}
#turn vector into a dataframe for analysis
rand_pun_.frame_2022G2 <- data.frame(random_times_pun_followed)








#make all column names consistent
colnames(rand_retal_.frame)[1] <- 'count_of_rule_follows'
colnames(rand_retal_.frame2022G1)[1] <- 'count_of_rule_follows'
colnames(rand_retal_.frame2022G2)[1] <- 'count_of_rule_follows'
colnames(rand_pile_.frame)[1] <- 'count_of_rule_follows'
colnames(rand_pile_.frame_2022G1)[1] <- 'count_of_rule_follows'
colnames(rand_pile_.frame_2022G2)[1] <- 'count_of_rule_follows'
colnames(rand_pass_.frame)[1] <- 'count_of_rule_follows'
colnames(rand_pass_.frame_2022G1)[1] <- 'count_of_rule_follows'
colnames(rand_pass_.frame_2022G2)[1] <- 'count_of_rule_follows'
colnames(rand_opp_.frame)[1] <- 'count_of_rule_follows'
colnames(rand_opp_.frame_2022G1)[1] <- 'count_of_rule_follows'
colnames(rand_opp_.frame_2022G2)[1] <- 'count_of_rule_follows'
colnames(rand_pun_.frame)[1] <- 'count_of_rule_follows'
colnames(rand_pun_.frame_2022G1)[1] <- 'count_of_rule_follows'
colnames(rand_pun_.frame_2022G2)[1] <- 'count_of_rule_follows'






#bind all dfs containing the counts of motif execution from the randomized reference models. bind possible grouping factors
tdr_count_rand_comb <- rbind(
  cbind(rand_retal_.frame, Group = "Retalialtion Group 1", groupid = "Group 1", rule = "Retaliation"),
  cbind(rand_retal_.frame2022G1, Group = "Retaliation Group 2", groupid = "Group 2", rule = "Retaliation"),
  cbind(rand_retal_.frame2022G2, Group = "Retaliation Group 3", groupid = "Group 3", rule = "Retaliation"),
  cbind(rand_pile_.frame, Group = "Pile-on Group 1", groupid = "Group 1", rule = "Pile-on"),  
  cbind(rand_pile_.frame_2022G1, Group = "Pile-on Group 2", groupid = "Group 2", rule = "Pile-on"),
  cbind(rand_pile_.frame_2022G2, Group = "Pile-on Group 3", groupid = "Group 3", rule = "Pile-on"),
  cbind(rand_pass_.frame, Group = "Pass-along Group 1", groupid = "Group 1", rule = "Pass-along"),
  cbind(rand_pass_.frame_2022G1, Group = "Pass-along Group 2", groupid = "Group 2", rule = "Pass-along"),
  cbind(rand_pass_.frame_2022G2, Group = "Pass-along Group 3", groupid = "Group 3", rule = "Pass-along"),
  cbind(rand_opp_.frame, Group = "Opportunism Group 1", groupid = "Group 1", rule = "Opportunism"),
  cbind(rand_opp_.frame_2022G1, Group = "Opportunism Group 2", groupid = "Group 2", rule = "Opportunism"),
  cbind(rand_opp_.frame_2022G2, Group = "Opportunism Group 3", groupid = "Group 3", rule = "Opportunism"),
  cbind(rand_pun_.frame, Group = "Punishment Group 1", groupid = "Group 1", rule = "Punishment"),
  cbind(rand_pun_.frame_2022G1, Group = "Punishment Group 2", groupid = "Group 2", rule = "Punishment"),
  cbind(rand_pun_.frame_2022G2, Group = "Punishment Group 3", groupid = "Group 3", rule = "Punishment")

)

#write.csv(tdr_count_rand_comb,  "tdr_count_rand_comb_global_75_perc.csv", row.names = FALSE)




#make all column names consistent
colnames(retal_real.frame)[1] <- 'count_of_rule_follows'
colnames(retal_real.frame2022G1)[1] <- 'count_of_rule_follows'
colnames(retal_real.frame2022G2)[1] <- 'count_of_rule_follows'
colnames(pile_real.frame_2021)[1] <- 'count_of_rule_follows'
colnames(pile_real.frame_2022G1)[1] <- 'count_of_rule_follows'
colnames(pile_real.frame_2022G2)[1] <- 'count_of_rule_follows'
colnames(pass_along_real_frame_2021)[1] <- 'count_of_rule_follows'
colnames(pass_along_real_2022G1_frame)[1] <- 'count_of_rule_follows'
colnames(pass_along_real_2022G2_frame)[1] <- 'count_of_rule_follows'
colnames(opp_real_2021_frame)[1] <- 'count_of_rule_follows'
colnames(opp_real_2022G1_frame)[1] <- 'count_of_rule_follows'
colnames(opp_real_2022G2_frame)[1] <- 'count_of_rule_follows'
colnames(punishment_real_2021_frame)[1] <- 'count_of_rule_follows'
colnames(punishment_real_2022G1_frame)[1] <- 'count_of_rule_follows'
colnames(punishment_real_2022G2_frame)[1] <- 'count_of_rule_follows'

# #bind all dfs containing the counts of motif execution from observed data. bind possible grouping factors
tdr_count_real_comb<- rbind(
  cbind(retal_real.frame, Group = "Retalialtion Group 1", groupid = "Group 1", rule = "Retaliation"),
  cbind(retal_real.frame2022G1, Group = "Retaliation Group 2", groupid = "Group 2", rule = "Retaliation"),
  cbind(retal_real.frame2022G2, Group = "Retaliation Group 3", groupid = "Group 3", rule = "Retaliation"),
  cbind(pile_real.frame_2021, Group = "Pile-on Group 1", groupid = "Group 1", rule = "Pile-on"),  
  cbind(pile_real.frame_2022G1, Group = "Pile-on Group 2", groupid = "Group 2", rule = "Pile-on"),
  cbind(pile_real.frame_2022G2, Group = "Pile-on Group 3", groupid = "Group 3", rule = "Pile-on"),
  cbind(pass_along_real_frame_2021, Group = "Pass-along Group 1", groupid = "Group 1", rule = "Pass-along"),
  cbind(pass_along_real_2022G1_frame, Group = "Pass-along Group 2", groupid = "Group 2", rule = "Pass-along"),
  cbind(pass_along_real_2022G2_frame, Group = "Pass-along Group 3", groupid = "Group 3", rule = "Pass-along"),
  cbind(opp_real_2021_frame, Group = "Opportunism Group 1", groupid = "Group 1", rule = "Opportunism"),
  cbind(opp_real_2022G1_frame, Group = "Opportunism Group 2", groupid = "Group 2", rule = "Opportunism"),
  cbind(opp_real_2022G2_frame, Group = "Opportunism Group 3", groupid = "Group 3", rule = "Opportunism"),
  cbind(punishment_real_2021_frame, Group = "Punishment Group 1", groupid = "Group 1", rule = "Punishment"),
  cbind(punishment_real_2022G1_frame, Group = "Punishment Group 2", groupid = "Group 2", rule = "Punishment"),
  cbind(punishment_real_2022G2_frame, Group = "Punishment Group 3", groupid = "Group 3", rule = "Punishment")

)

#write.csv(tdr_count_real_comb,  "tdr_count_real_comb_global_75_perc.csv", row.names = FALSE)



```



##Plot observed motif count compared to random violin and barplots
```{r}

#load randomized df
tdr_count_rand_comb <- read_csv("tdr_count_rand_comb_global_75_perc.csv")


#calculate means of randomized motif execution for each rule
tdr_count_rand_comb_mean <- tdr_count_rand_comb  %>% group_by(groupid, rule) %>% summarise(mean_value = mean(count_of_rule_follows, na.rm = TRUE))





#load observed df
tdr_count_real_comb <- read_csv("tdr_count_real_comb_global_75_perc.csv")



#set order in which rules should be displayed in the plot
custom_order <- c("Retaliation","Pile-on",  "Pass-along", "Opportunism", "Punishment")
tdr_count_rand_comb$rule <- factor(tdr_count_rand_comb$rule, levels = custom_order)

tdr_count_real_comb$rule <- factor(tdr_count_real_comb$rule, levels = custom_order)

#compressed violin version

rule_order <- c("Retaliation","Pile-on","Pass-along","Opportunism","Punishment")
# Use only rules that actually exist in data just in case but keep the specified order
present_rules <- intersect(rule_order, unique(tdr_count_rand_comb$rule))
groups <- sort(unique(tdr_count_rand_comb$groupid))

#Build axis order with rules clustered together and then g1 g2 g3 of each 
#makes it so goes through groups first then rules
#looks like Retaliation – G1,G2,G3 | Pile-on – G1,G2,G3 | ...
axis_order <- expand.grid(
  groupid = groups,
  rule    = present_rules,
  KEEP.OUT.ATTRS = FALSE
) |>
  mutate(xcat = paste0(rule, "–", groupid)) |>
  pull(xcat)
#make sure only pulling out appropriate rules and groups

tdr_rand2 <- tdr_count_rand_comb |>
  filter(rule %in% present_rules, groupid %in% groups) |>
  mutate(
    rule = factor(rule, levels = present_rules),
    xcat = factor(paste0(rule, "–", groupid), levels = axis_order)
  )
#same thing for observed motif execution
tdr_real2 <- tdr_count_real_comb |>
  filter(rule %in% present_rules, groupid %in% groups) |>
  mutate(
    rule = factor(rule, levels = present_rules),
    xcat = factor(paste0(rule, "–", groupid), levels = axis_order)
  )

#create separator positions between rule blocks
g_per_rule    <- length(groups)
sep_positions <- (seq_len(length(present_rules) - 1) * g_per_rule) + 0.5

rule_violin <- ggplot(tdr_rand2, aes(x = xcat, y = count_of_rule_follows)) +
#separators behind violins
geom_vline(xintercept = sep_positions, linetype = "dashed",
             color = "grey60", linewidth = 0.8) +
#randomized reference as gray violins
geom_violin(fill = "grey70", color = "black", alpha = 1,
              width = 0.9, trim = FALSE, scale = "width") +
#mean bar across each violin
   stat_summary(fun = mean, geom = "crossbar", width = 0.6,
              color = "black", linewidth = 0.3)+
#observed value as a red horizontal line across each category
  geom_segment(
    data = tdr_real2,
    aes(x = as.numeric(xcat) - 0.35,
        xend = as.numeric(xcat) + 0.35,
        y = count_of_rule_follows,
        yend = count_of_rule_follows),
    inherit.aes = FALSE,
    linewidth = 1.2, color = "red"
  ) +
  labs(
    x = NULL,
    y = "Count of motif execution", 
    title = "Observed vs. randomized reference by rule and group (violin)"
  ) +
  theme_minimal(base_size = 20) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.8),
    axis.text.x  = element_text(angle = 45, hjust = 1),
    axis.ticks.x = element_line()
  )

rule_violin
#15 by 12 or 12 by 15









#short bar plot to attach
head(tdr_count_rand_comb_mean)
head(tdr_count_real_comb)


tdr_count_both_comb <- tdr_count_rand_comb_mean %>%
  left_join(tdr_count_real_comb, by = c("groupid", "rule"))
#obtain the difference between observed count and the mean of the reference distribution
tdr_count_both_comb$dif <- tdr_count_both_comb$count_of_rule_follows - tdr_count_both_comb$mean_value



rule_order <- c("Retaliation","Pile-on","Pass-along","Opportunism","Punishment")
# Use only rules that actually exist in data just in case but keep the specified order
present_rules <- intersect(rule_order, unique(tdr_count_both_comb$rule))
groups <- sort(unique(tdr_count_both_comb$groupid))

#Build axis order with rules clustered together and then g1 g2 g3 of each 
#makes it so goes through groups first then rules
#looks like Retaliation – G1,G2,G3 | Pile-on – G1,G2,G3 | ...
axis_order <- expand.grid(
  groupid = groups,
  rule    = present_rules,
  KEEP.OUT.ATTRS = FALSE
) |>
  mutate(xcat = paste0(rule, "–", groupid)) |>
  pull(xcat)
#make sure only pulling out appropriate rules and groups

tdr_count_both_comb2 <- tdr_count_both_comb |>
  filter(rule %in% present_rules, groupid %in% groups) |>
  mutate(
    rule = factor(rule, levels = present_rules),
    xcat = factor(paste0(rule, "–", groupid), levels = axis_order)
  )

#create separator positions between rule blocks
g_per_rule    <- length(groups)
sep_positions <- (seq_len(length(present_rules) - 1) * g_per_rule) + 0.5



rule_bar <- ggplot(tdr_count_both_comb2, aes(x = xcat, y = dif)) +
  #separators behind bars
  geom_vline(xintercept = sep_positions, linetype = "dashed",
             color = "grey60", linewidth = 0.8) +
  #bars from supplied y
  geom_col(fill = "grey60", color = "black") +
  labs(
    x = NULL,
    y = "Count of motif execution",
    title = "Observed vs. randomized reference by rule and group (bar)"
  ) +
  theme_minimal(base_size = 20) +
  scale_y_continuous(limits = c(-450, 100)) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.8),
    axis.text.x  = element_text(angle = 90, hjust = 1),
    axis.ticks.x = element_line()
  )

rule_bar





#keep the top plot as is visually, just hide its x label ticks so they don't repeat
rule_violin_top <- rule_violin +
  theme(
    axis.title.x = element_blank(),
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    plot.margin  = margin(5.5, 5.5, 0, 5.5) 
  )

#keep bar plot exactly as is but tweak bottom margin
rule_bar_bottom <- rule_bar +
  theme(
    plot.margin = margin(0, 5.5, 5.5, 5.5)
  )

#stack two plots together. The violins on top (bigger), bars below (smaller)
combined <- rule_violin_top / rule_bar_bottom +
  plot_layout(heights = c(2, 1), guides = "keep")

combined
#12 by 10

#P values added manually in inkscape for easier formatting

```




